---
title: "SepTest"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

**SepTest** provides statistical tools for testing _first-order separability_ in spatio-temporal point processes, that is, whether a spatio-temporal intensity function can be expressed as the product of purely spatial and purely temporal components.

The package implements several hypothesis testing procedures, including exact and asymptotic methods for both Poisson and non-Poisson processes. Available methods include global envelope tests, chi-squared-type statistics, and a novel Hilbert–Schmidt Independence Criterion (HSIC)–based test using block or pure permutation schemes.

The package covers the simulation studies and applications presented in Ghorbani et al. (2021, 2025).
 
## Installing the package

You can install the development version of SepTest from GitHub using:

```{r}
# Install remotes if not already installed
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes", repos = "https://cloud.r-project.org")
}
remotes::install_github("mghorbani01/SepTest")
```

## Examples
```{r}
library(SepTest)

if (requireNamespace("dHSIC", quietly = TRUE)) {

  set.seed(123)

  # Simulated spatio-temporal data
  X <- cbind(
    runif(100),          # spatial x-coordinate
    runif(100),          # spatial y-coordinate
    runif(100, 0, 10)    # temporal component
  )

  # Pure permutation HSIC test
  result <- dHS.test(
    sim.procedure = "pure_per",
    X = X,
    nsim = 199,
    bandwidth = 0.05
  )
  print(result$p.value)

  # Block permutation HSIC test
  result_block <- dHS.test(
    sim.procedure = "block_per",
    X = X,
    nblocks = 5,
    nperm = 100,
    bandwidth = 0.05
  )
  print(result_block$p.value.bw)
}
```

## Documentation

For full documentation of each function, use:

```{r}
help(package = "SepTest")
```
## Bug Reports

Please report bugs or feature requests at:

https://github.com/mghorbani01/SepTest/issues

When reporting issues, please include a minimal reproducible example and your
sessionInfo() output.

## License

GPL-3
